### 2024-05-18 manuscripts/ilr-review copy/manuscript_main.tex
create new table

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main Text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}

@@

Using the CPS from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes. In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings. I also study the effect of identifying as Hispanic on labor market outcomes.

I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences. I find that inter-ethnic children with Hispanic sounding last name earn 5 percentage points less than those without a Hispanic last name. The gap becomes statistically insignificant after controlling for education. Thus, I do not find a significant effect of having a Hispanic last name. Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.

The rest of the paper is organized as follows. Section \ref{sec:data} provides an overview of the data and summary statistics of the sample. Section \ref{sec:emp_model} introduces the empirical framework. I report the results in section \ref{sec:results}. Finally, I offer a brief conclusion in section \ref{sec:con1}.

\section{Data}\label{sec:data}

@@

\subsection{Children of the four parental types}

@@

% On the one hand, WW children are more educated, with approximately 14 years of education for men and women. WW children have an unemployment rate of 4\% for men and women. They have log hourly earnings of 2.51 for men and 2.32 for women. Also, their annual log earnings are 10.29 for men and 9.43 for women.

% On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively. HH men and women have an unemployment rate of 7\% and 6\%. Their annual log earnings are 10.01 and 9.53 for men and women, respectively.

% Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children. Also, WH and HW are comparable, unlike WW and HH children. These two groups are important to the analysis. Children of HW are going to be the treatment group. HW children have a Hispanic father with a Hispanic last names. At the same time, the WH children have White fathers with a non-Hispanic, or White, last names. Summary statistics of these two groups, Table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.

% WH children have an average of 13.58 and 13.79 years of education for men and women. The unemployment rate for men and women is 5\%. HW men and women obtained 13.22 and 13.42 years of schooling. Men's unemployment rate is 7\% and women's 6\%.

% In panel C of Table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic. \footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic. There is considerable attrition in Hispanic identification of inter-ethnic children. Among WH children, 74\% of men and 78\% of women identify as Hispanic. Among HW children, 83\% of men and 82\% of women identify as Hispanic. These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \autocite{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}. Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic. These 5\% of WW children are most likely third generation--- or higher--- children of immigrants. For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.

% In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6). Overall, HH children do worse than WW children. HH has one year less of education compared to WW, and are 2 percentage points more likely to be unemployed. HH men earn 9\% less than WW men, and HH women make 2\% less than WW women. Therefore, these two groups, HH and WW, are different on many observable qualities.

% The differentials between WH and HW are not as severe as the ones between HH and WW. HW children are still worse off than WH but are much better than HH. HW men attain 0.36 fewer years of education than WH men. HW women achieve 0.37 fewer years of schooling than WH women. HW children are 1 percentage points more likely to be unemployed. HW men earn 1\% less annual earnings as WH men, and HW women earn 3\% less. These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.

% In Table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH. Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists. The one difference is that those who identify as Hispanic do slightly worse.

\subsection{Synthetic parents}

@@

\section{Empirical Strategy}\label{sec:emp_model}

@@

\subsection{Estimating the effect of having a Hispanic last name}

@@

\subsection{Estimating the effect of identifying as Hispanic}

@@

\subsection{Threats to Identification}

@@

\section{From the Data: The Differences Between HW and WH Couples}\label{sec:hw-wh-couples-data}

@@

\section{Results}\label{sec:results}

@@

\subsection{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}

@@

\section{Conclusion}\label{sec:con1}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main Text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Using the CPS from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes. In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings. I also study the effect of identifying as Hispanic on labor market outcomes.

% I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences. I find that inter-ethnic children with Hispanic sounding last name earn 5 percentage points less than those without a Hispanic last name. The gap becomes statistically insignificant after controlling for education. Thus, I do not find a significant effect of having a Hispanic last name. Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.

\section*{Data}\label{sec:data}

@@

\subsection*{Children of the four parental types}

@@

\subsection*{Synthetic parents}

@@

\section*{Empirical Strategy}\label{sec:emp_model}

@@

\subsection*{Estimating the effect of having a Hispanic last name}

@@

\subsection*{Estimating the effect of identifying as Hispanic}

@@

\subsection*{Threats to Identification}

@@

\section*{From the Data: The Differences Between HW and WH Couples}\label{sec:hw-wh-couples-data}

@@

\section*{Results}\label{sec:results}

@@

\subsection*{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}

@@

\section*{Conclusion}\label{sec:con1}
